Highly Available Transactions: Virtues and Limitations

### 2013-03-19 writeup/hat-analysis.tex
more writing

@@

Let's sketch out a few conditions, then we'll discuss them later:
* Lost Update
	* Simple proof: T1 R(x) W(x=x+1) T2 R(x) W(x=x+1); execute on opposite sides 
* Read Skew
	* Similar easy proof
* Write Skew
	* Similar easy proof
* Anti-dependency cycles
	* Serializability
* Recency guarantees on data items
	* CAP

>>>>

Read your writes is unavailable in a highly available system. Consider
a client that executes the following two transactions in succession:
\vspace{-.5em}
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: r_x(a)
\end{align*}
If the client executes $T_1$ against a server that is partitioned from
the rest of the other servers, the server must allow $T_1$ to
commit. If the client executes $T_2$ against the same (partitioned)
server, then it will be able to read its writes. However, if the
network topology shifts and the client can only contact a different
server which is partitioned from the rest of the cluster, then the
client will not be able to read its own writes and the system will
have to either stall indefinitely to allow the client to read her
writes (violating transactional availability) or will have to
sacrifice read your writes. Accordingly, read your writes requires
stickiness, and, because causal consistency requires read your writes,
causal consistency also requires stickiness.

Read-your-writes can be accomplished via client-side caching and
alternatively comes ``for free'' when requests are routed to a sticky
set of servers. Achieving causal consistency requires maintaining
read-your-writes along with the prior session guarantees. Both are
reasonable strategies often employed in real-world
systems~\cite{vogels-defs}, albeit at the cost of true high
availability or client-side storage.

We discuss additional uses for stickiness in Section~\ref{sec:discussion}.

\subsubsection{Additional Guarantees}

A HAT system can make \textbf{limited consistency} guarantees. It can
execute commutative and logically monotonic~\cite{needed} operations
without the risk of invalidating (also monotonic) application-level
integrity constraints. Our goal in this paper is not to sketch the
entire space of consistency models that are achievable (see
Section~\ref{sec:futurework}, however we specifically evaluate TPC-C
transaction semantics under HAT consistency guarantees in
Section~\ref{sec:Evaluation}.

As we briefly discussed in Section~\ref{sec:availabile}, a client
requiring \textbf{durability} of surviving $F$ server faults requires
at least $F+1$-availability.

Finally, to require that servers propagate writes between replicas, we
can require \textbf{convergence}, or eventual consistency for each
data item: in the absence of new mutations to a data item, in the
absence of partitions, all servers eventually agree on the value for
each item. This is typically accomplished by any number of
anti-entropy protocols, which periodically update neighboring servers
with the latest value for each data item.

While there are infinitely many HAT models
(Section~\ref{sec:futurework}), at this point, we have exhausted our
range of achievable, previously defined (useful) HAT semantics. Before
summarizing our possibility results, we will present impossibility
results for HATs, also defined in terms of previously identified isolation and consistency anomalies.

\subsubsection{Unachievable ACID Isolation}

* Lost Update
	* Simple proof: T1 R(x) W(x=x+1) T2 R(x) W(x=x+1); execute on opposite sides 
* Read Skew
	* Similar easy proof
* Write Skew
	* Similar easy proof
* Anti-dependency cycles
	* Serializability
* Recency guarantees on data items
	* CAP

\subsection{Summary}
\label{sec:hat-summary}

 \newcommand{\lostupdate}{$^\dagger$}
 \newcommand{\rwskew}{$^\ddagger$}
 \newcommand{\mphantom}{$^\star$}
 \newcommand{\linearizable}{$^\oplus$}

\begin{table}
\begin{tabular}{| c | p{6cm} | }\hline
HA & Transactional Atomicity (TA), Read Uncommitted (RU), Read
Committed (RC), Cut Isolation (CI), Monotonic Reads (MR), Monotonic
Writes (MW), Writes Follow Reads (WFR)\\\hline Sticky-HA & Read Your
Writes (RYW), Causal Consistency\\\hline Unavailable & Cursor
Stability (CS)\lostupdate, Read Consistency (CR)\lostupdate, Snapshot
Isolation (SI)\lostupdate, Repeatable Read (RR)\lostupdate\rwskew,
Serializability (S)\lostupdate\rwskew\mphantom, Causal
Serializability\lostupdate\rwskew\mphantom, Strict
Serializability\lostupdate\rwskew\mphantom\linearizable \\\hline
\end{tabular}
\caption{Summary of highly available, sticky highly available, and
  unavailable models considered in this paper. Unavailable models are
  labeled by cause of unavailability: preventing lost
  update\lostupdate, preventing read/write skew\rwskew, preventing
  phantoms\mphantom, and requiring linearizability\linearizable.}
\label{table:hatcompared}
\end{table}

\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8]
  \tikzstyle{sticky}=[rectangle,draw=blue!50,fill=blue!20,thick]
  \tikzstyle{noha}=[ellipse,draw=red!50,fill=red!20,thick, inner sep=0pt,minimum size=12pt]

  \tikzstyle{every node}=[font=\small]

 \node[draw=none,fill=none] (ci) at (1.2, 1.6) {\underline{CI}};
 \node[draw=none,fill=none] (rc) at (1.2, .8) {\underline{RC}};
 \node[draw=none,fill=none] (ru) at (1.2, 0) {\underline{RU}};

 \node[draw=none,fill=none] (ta) at (0, 0) {\underline{TA}};
 \node[draw=none,fill=none] (mr) at (2.4, 0) {\underline{MR}};
 \node[draw=none,fill=none] (mw) at (3.6, 0) {\underline{MW}};
 \node[draw=none,fill=none] (wfr) at (4.8,0) {\underline{WFR}};
 \node at (6,0) [sticky] (ryw) {RYW};

 \node at (3.6, 2) [sticky] (causal) {Causal};
 \node[noha] (cs) at (1.7, 2.4) {CS};
 \node[noha] (rr) at (.7, 2.6) {RR};
 \node[noha] (si) at (-.4, 2.2) {SI};
 \node[noha] (cr) at (-.4, 1.2) {CR};
 \node[noha] (s) at (.5, 3.5) {S};
 \node[noha] (causal-s) at (3.6, 3.5) {Causal S};
 \node[noha] (strict-s) at (6, 3.5) {Strict S};

 \draw [->] (ru) -- (rc);
 \draw [->] (rc) -- (ci);

 \draw [->, blue] (mr) -- (causal);
 \draw [->, blue] (mw) -- (causal);
 \draw [->, blue] (wfr) -- (causal);
 \draw [->, blue] (ryw) -- (causal);

 \draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ru) -- (mr);
 \draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (rc) -- (ta);
 \draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ci) -- (ta);
 \draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (rc) -- (mr);
 \draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ci) -- (mr);
 \draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ta) -- (ru);
 \draw[snake=coil, blue, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (ru) -- (causal);
 \draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (mr) -- (mw);
 \draw[snake=coil, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (wfr) -- (mw);
 \draw[snake=coil, blue, segment aspect=0, segment amplitude=.75pt, segment length=2pt] (wfr) -- (ryw);

 \draw [->, red] (ci) -- (cs);
 \draw [->, red] (cs) -- (rr);
 \draw [->, red] (ci) -- (si);
 \draw [->, red] (rr) -- (s);
 \draw [->, red] (si) -- (s);
 \draw [->, red] (ta) -- (cr);
 \draw [->, red] (rc) -- (cr);
 \draw [->, red] (cr) -- (si);
 \draw [->, red] (ta) -- (rr);
 \draw [->, red] (s) -- (causal-s);
 \draw [->, red] (causal) -- (causal-s);
 \draw [->, red] (causal-s) -- (strict-s);

\end{tikzpicture}
\label{fig:hat-order}
\caption{HAT, sticky HAT (in boxes), and unavailable models (circled)
  from Table~\protect\ref{table:hatcompared} compared graphically. Directed
  edges represent ordering by model strength, squiggle edges represent
  (transitive) compatible models, while absence of edges represents
  incomparable models. The combination of all underlined models
  represents maximum ACID HAT semantics, while the combination of
  underlined and boxed models represents maximum ACID sticky HAT
  models.}
\label{fig:hatcompared}
\end{figure}

\subsection{Discussion}
\label{sec:discussion}

@@

Let's sketch out a few conditions, then we'll discuss them later:
